---
title: "CD8 BCR"
output: html_document
date: "2023-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

#----R load----
### Load R packages
```{r}
library(tidyverse)
library(pheatmap)
```

#----Data load----




####Load IMGT

```{r}
# https://www.imgt.org/IMGTrepertoire/index.php?section=LocusGenes&repertoire=GeneOrder&species=human&group=IGK
IGHVIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGHV.txt',sep="\t")
IGHJIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGKJ.txt',sep="\t")
IGKVIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGKV.txt',sep="\t")
IGLJIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGLJ.txt',sep="\t")
IGLKIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGLK.txt',sep="\t")
IGLVIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGLV.txt',sep="\t")
IGHDIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/BCRgene/IGHD.csv',sep=",")
IGKJIMGT=data.frame(IMGT_gene_name=paste0("IGKJ",1:5),order=seq(77,81))
```

#### Load obs

```{r}
BCRobs=read.csv("/data5/maolp/Gaofeng_All_matrix/Allcount/All_scanpyData/Data/BCRobs.csv")
BCRobs
# BCRobs<-subset(BCRobs,Last_cell_type=="Naïve CD8 T")
```

```{r}
# subBCRK=AllBCR[AllBCR$chain=="BCRK",]
# subBCRH=AllBCR[AllBCR$chain=="BCRH",]
```


```{r}
subBCRHobs<-BCRobs
```
```{r}
colnames(subBCRHobs)[grep("call",colnames(subBCRHobs))] %>% .[grep("_1",.)]
```




```{r}
# subBCRHobs[,c("cdr3","IR_VDJ_1_junction_aa")]
# subBCRHobs2<-subBCRHobs[!duplicated(subBCRHobs$barcodename),]
subBCRHobs2=subBCRHobs
```

### Get BCRB

```{r}
subVDJ=subBCRHobs2[,c("AdjustedID","New_Body",colnames(subBCRHobs2)[grep("call",colnames(subBCRHobs))] %>% .[grep("_1",.)])]
# subVDJ$Week=
```
### Get BCRK
```{r}
subBCRK=subBCRHobs2[,c("AdjustedID","New_Body",colnames(subBCRHobs2)[grep("call",colnames(subBCRHobs))]
                     %>% .[grep("_1",.)])] 
# subBCRK=
```




```{r}
unique(subVDJ$AdjustedID)
subVDJ$Week=substring(subVDJ$AdjustedID,2,5) %>% as.numeric()
subVDJ$Satge<-ifelse(subVDJ$Week>26,"Late","Early")
subVDJ$OrganStage<-paste0(subVDJ$Satge,"_",subVDJ$New_Body)
```

```{r}
subVDJ$BCRHVJ=paste0(subVDJ$IR_VDJ_1_v_call,"_",subVDJ$IR_VDJ_1_j_call)
```

```{r}
table(subVDJ$AdjustedID,subVDJ$BCRHVJ) %>% as.data.frame()
```



#----All BCR Freq----


### All percentage heatmap

```{r}
subBCRHh1=subVDJ[grep("B",subVDJ$AdjustedID),]
```

```{r}
subBCRHh1 = subVDJ[grep("B", subVDJ$AdjustedID), ]
BCRHVJ = table(subBCRHh1$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
BCRHVJ$Per = BCRHVJ$Freq / sum(BCRHVJ$Freq) * 100
PBMCBCRHVJ_separated <- BCRHVJ %>%
  separate(Var1, into = c("Vgene", "Jgene"), sep = "_")

PBMCBCRHVJ_wide <- PBMCBCRHVJ_separated %>% select(-Freq) %>%
  spread(key = Vgene, value = Per)
rownames(PBMCBCRHVJ_wide) <- PBMCBCRHVJ_wide$Jgene
PBMCBCRHVJ_wide <- PBMCBCRHVJ_wide[, -1]
PBMCBCRHVJ_wide[is.na(PBMCBCRHVJ_wide)] <- 0
library(pheatmap)
pheatmap(
  PBMCBCRHVJ_wide,
  border_color = "white",
  color = rev(colorRampPalette(
    c('#C6403D', "#f6c619", 'white', '#246BAE')
  )(100))
)
```


### All BCR VDJheatmap

#### Main Function

##### BCRHIMGTsub
```{r}
BCRHIMGTsub=IGHVIMGT[IGHVIMGT$IMGTgenename %in% subVDJ$IR_VDJ_1_v_call,] %>% arrange(IMGTgeneorder)
colnames(BCRHIMGTsub)<-c("IMGT_gene_name","IMGTgeneorder")
```


```{r}
library(tidyverse)
library(pheatmap)

VDJheatmap <- function(data,pattern, BCRHinIMGT=BCRHIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRHVJ = table(sub_data$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRHVJ$Per = BCRHVJ$Freq / sum(BCRHVJ$Freq)*100
  
  separated_data <- BCRHVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
missing_cols <- setdiff(BCRHinIMGT$IMGT_gene_name, colnames(wide_data))


for(col in missing_cols) {
  wide_data[[col]] <- 0
}
  wide_data<-wide_data[,BCRHinIMGT$IMGT_gene_name]
  pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c('#C6403D', "#f6c619","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  return(wide_data)
}
```


```{r}
library(tidyverse)
library(pheatmap)

VDJheatmap_data <- function(data,pattern, BCRHinIMGT=BCRHIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRHVJ = table(sub_data$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRHVJ$Per = BCRHVJ$Freq / sum(BCRHVJ$Freq)
  
  separated_data <- BCRHVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  return( wide_data)
  # wide_data<-wide_data[,BCRHinIMGT$IMGT_gene_name]
  # pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c('#C6403D', "#f6c619","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  # return(wide_data)
}
```

```{r}
IGHVdata=subVDJ
sub_IGHVdata = IGHVdata[grep("Early_PBMC", IGHVdata[["OrganStage"]]),]
BCRHVJ = table(sub_IGHVdata$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
BCRHVJ$Per = BCRHVJ$Freq / sum(BCRHVJ$Freq)

separated_IGHVdata <- BCRHVJ %>%
  separate(Var1, into = c("Vgene", "Jgene"), sep = "_")

wide_IGHVdata <- separated_IGHVdata %>% select(-Freq) %>%
  spread(key = Vgene, value = Per)

rownames(wide_IGHVdata) <- wide_IGHVdata$Jgene
wide_IGHVdata <- wide_IGHVdata[, -1]
wide_IGHVdata[is.na(wide_IGHVdata)] <- 0
missing_cols <- setdiff(BCRHIMGTsub$IMGT_gene_name, colnames(wide_IGHVdata))
# return( wide_IGHVdata)
```


#### Early_PBMC
```{r}
VDJheatmap(subVDJ,pattern =  "Early_PBMC",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_PBMC")
```


#### Late_PBMC no cluster
```{r}
VDJheatmap(subVDJ, pattern = "Late_PBMC",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Late_PBMC")
```

#### Late_PBMC cluster
```{r}
VDJheatmap(subVDJ, pattern = "Late_PBMC",id_col="OrganStage",cluster_col=T,cluster_row=T,title = "Late_PBMC")
```
####  Early_PBMC cluster
```{r}
VDJheatmap(subVDJ,pattern =  "Early_PBMC",id_col="OrganStage",cluster_col=T,cluster_row=,title = "Early_PBMC")
```

####  All_PBMC cluster
```{r}
VDJheatmap(subVDJ, "B",cluster_col=F,cluster_row=F,title = "PBMC")
```

#### Early_Liver
```{r}
VDJheatmap(subVDJ,pattern = "Early_Liver",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Liver")
```

#### Early_Spleen
```{r}
VDJheatmap(subVDJ,pattern = "Early_Spleen",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Spleen")
```
#### Early_Thymus
```{r}
VDJheatmap(subVDJ,pattern = "Early_Thymus",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Thymus")
```


#----Per Sample BCRH----

### Get Sample percentage
```{r}
subBCRHh1=subVDJ[grep("B",subVDJ$AdjustedID),]
Sample_BCRHVJ=table(subBCRHh1$AdjustedID,subBCRHh1$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_BCRHVJ)<-c("AdjustedID","VDJ","Freq")

total_freq_by_id <- Sample_BCRHVJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_BCRHVJ <- left_join(Sample_BCRHVJ, total_freq_by_id, by = "AdjustedID")


Sample_BCRHVJ <- Sample_BCRHVJ %>% 
  mutate(Per = Freq / total_freq*100)

Sample_BCRHVJ <- Sample_BCRHVJ %>% arrange(AdjustedID)
Sample_BCRHVJ$Week=substring(Sample_BCRHVJ$AdjustedID,2,5) %>% as.numeric()
```
```{r}
subBCRK
```




```{r}
unique(Sample_BCRHVJ$AdjustedID)
```
### Sample top BCRH

```{r}

library(dplyr)
library(ggplot2)

Sample_BCRHVJ$Week <- as.numeric(Sample_BCRHVJ$Week)
Sample_BCRHVJ$Per <- as.numeric(Sample_BCRHVJ$Per)


top_VDJ <- Sample_BCRHVJ %>% 
  group_by(VDJ) %>%
  summarise(Total_Per = sum(Per)) %>%
  top_n(10, Total_Per) %>%
  pull(VDJ)

Sample_BCRHVJ <- Sample_BCRHVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% top_VDJ, VDJ, "Other"))
Sample_BCRHVJ$VDJ_Top10<-factor(Sample_BCRHVJ$VDJ_Top10,levels = c(as.character(top_VDJ),"Other"))
```




```{r}
table(Sample_BCRHVJ$AdjustedID)
```

#### Plot: top BCRH line 

```{r}

ggplot(Sample_BCRHVJ, aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  geom_line() +
  scale_color_manual(values = c(rainbow(10), "gray")) +
  labs(x = "Week", y = "Per", color = "VDJ") +
  theme_classic()
```
### PerSample Cor BCRHV
```{r}

PBMCBCRH_cor_results <- Sample_BCRHVJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
PBMCBCRH_cor_results<-PBMCBCRH_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(PBMCBCRH_cor_results)
subset(PBMCBCRH_cor_results,p.adjusted<0.1) %>% arrange(cor)
PBMCBCRH_cor_results$VDJcor<-ifelse(PBMCBCRH_cor_results$p.value>0.05,"NA",ifelse(PBMCBCRH_cor_results$cor>0,"+","-"))
```
```{r}
subset(PBMCBCRH_cor_results,p.adjusted<0.2)
```
#### Plot: top sig BCRH line 



```{r}
top_VDJ=subset(PBMCBCRH_cor_results,p.adjusted<0.1) %>% arrange(cor) %>% .$VDJ %>% .[-1]
Sample_BCRHVJ <- Sample_BCRHVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% top_VDJ, VDJ, "Other")) %>% 
mutate(VDJ_Top10=factor(VDJ_Top10,levels = c(as.character(top_VDJ),"Other")))
subset(Sample_BCRHVJ,VDJ_Top10!= "Other") %>% 
ggplot() +
  geom_line( aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  scale_color_manual(values = c(rainbow(length(top_VDJ)), "gray")) +
  labs(x = "Week", y = "Percentage", color = "VDJ") +
  theme_classic()
```

```{r}
Sample_BCRHVJ %>% arrange(desc(Per))
```

```{r}
# 
# PBMCBCRHVJ_separated <- BCRHVJ%>%
#   separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
# 
# PBMCBCRHVJ_wide <- PBMCBCRHVJ_separated %>%select(-Freq) %>% 
#   spread(key =Vgene, value = Per)
# rownames(PBMCBCRHVJ_wide)<-PBMCBCRHVJ_wide$Jgene
# PBMCBCRHVJ_wide<-PBMCBCRHVJ_wide[,-1]
# PBMCBCRHVJ_wide[is.na(PBMCBCRHVJ_wide)]<-0
```
```{r}
# library(pheatmap)
# pheatmap(PBMCBCRHVJ_wide,border_color = "white",color =rev( colorRampPalette(c('#C6403D', 'white','#246BAE'))(100)))
```



```{r}
# intersect(BCRHIMGT$IMGT_gene_name,subVDJ$IR_VDJ_1_v_call)
```

```{r}
# setdiff(subVDJ$IR_VDJ_1_v_call,BCRHIMGT$IMGT_gene_name)
```





### BCRHfunction Data


```{r}
library(ggplot2)
library(reshape2) # for melt function

VDJfun <- function(data,pattern, BCRHinIMGT=BCRHIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRHVJ = table(sub_data$BCRHVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRHVJ$Per = BCRHVJ$Freq / sum(BCRHVJ$Freq)*100
  
  separated_data <- BCRHVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(BCRHinIMGT$IMGT_gene_name, colnames(wide_data))

  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  
  wide_data <- wide_data[,BCRHinIMGT$IMGT_gene_name]
  
  return(wide_data)
}
```
#---------------------------

### IMGT_HighV-QUEST
```{r}
library(data.table)
read_and_assign <- function(folder) {
  file_path <- paste0("../Figure5_BCRIMGT/IMGT_HighV-QUEST/",folder, "/6_Junction.txt")
  # print(file_path)
  assign(paste0(folder, "_data"), fread( file_path ,header = T), envir = .GlobalEnv)
}


folders <- c("IGH", "IGK", "IGL", "IGH", "IGH")


lapply(folders, read_and_assign)
```
```{r}
library(data.table)
read_and_assign <- function(folder) {
  file_path <- paste0("../Figure5_BCRIMGT/IMGT_HighV-QUEST/",folder, "/6_Junction.txt")
  # print(file_path)
  assign(paste0(folder, "_data"), fread( file_path ,header = T), envir = .GlobalEnv)
}


folders <- c("IGH", "IGK", "IGL", "IGH", "IGH")


lapply(folders, read_and_assign)
```

```{r}
dir("../../HFB0404TCR/Figure4_TCR_IMGTCDR3/IMGT_HighV-QUEST/")
```
```{r}
library(data.table)
read_and_assign <- function(folder) {
  file_path <- paste0("../../HFB0404TCR/Figure4_TCR_IMGTCDR3/IMGT_HighV-QUEST/",folder, "/6_Junction.txt")
  # print(file_path)
  assign(paste0(folder, "_data"), fread( file_path ,header = T), envir = .GlobalEnv)
}


folders <- c("TRA","TRB")


lapply(folders, read_and_assign)
```


```{r}
IGH_data
median(nchar(IGH_data$`N1-REGION`))
mean(nchar(IGH_data$`N1-REGION`))
```
```{r}
median(nchar(IGH_data$`N2-REGION`))
mean(nchar(IGH_data$`N2-REGION`))
```
```{r}
mean(nchar(IGH_data$`5'J-REGION`))
```
```{r}
mean(nchar(IGH_data$`3'V-REGION`))
```
```{r}
mean(nchar(IGH_data$`N-REGION`))
```
```{r}
mean(nchar(IGH_data$`D-REGION`))
```
```{r}
calculate_mean_length <- function(data, column_name) {
  if(!column_name %in% names(data)) {
    stop("Column not found in the data frame")
  }
  

  mean_length <- mean(nchar(data[[column_name]]))
  

  result <- data.frame(
    "Region" = column_name,
    "MeanLength" = mean_length
  )
  

  return(result)
}
```


```{r}
IGHmean_lengths <- rbind(
  calculate_mean_length(IGH_data, 'N1-REGION'),
  calculate_mean_length(IGH_data, 'N2-REGION'),
  calculate_mean_length(IGH_data, '5\'J-REGION'),
  calculate_mean_length(IGH_data, '3\'V-REGION'),
  calculate_mean_length(IGH_data, 'N-REGION'),
  calculate_mean_length(IGH_data, 'D-REGION')
)


print(IGHmean_lengths)
```
```{r}
TRAmean_lengths <- rbind(
  calculate_mean_length(TRA_data, 'N1-REGION'),
  calculate_mean_length(TRA_data, 'N2-REGION'),
  calculate_mean_length(TRA_data, '5\'J-REGION'),
  calculate_mean_length(TRA_data, '3\'V-REGION'),
  calculate_mean_length(TRA_data, 'N-REGION'),
  calculate_mean_length(TRA_data, 'D-REGION')
)


```

```{r}
TRBmean_lengths <- rbind(
  calculate_mean_length(TRB_data, 'N1-REGION'),
  calculate_mean_length(TRB_data, 'N2-REGION'),
  calculate_mean_length(TRB_data, '5\'J-REGION'),
  calculate_mean_length(TRB_data, '3\'V-REGION'),
  calculate_mean_length(TRB_data, 'N-REGION'),
  calculate_mean_length(TRB_data, 'D-REGION')
)



```


```{r}
IGKmean_lengths <- rbind(
  calculate_mean_length(IGK_data, 'N1-REGION'),
  calculate_mean_length(IGK_data, 'N2-REGION'),
  calculate_mean_length(IGK_data, '5\'J-REGION'),
  calculate_mean_length(IGK_data, '3\'V-REGION'),
  calculate_mean_length(IGK_data, 'N-REGION'),
  calculate_mean_length(IGK_data, 'D-REGION')
)
```
```{r}

data_list <- list(
  IGH= IGH_data,
  IGL = IGL_data,
  IGK = IGK_data,
 
  TRA = TRA_data,
  TRB = TRB_data
)

region_names <- c('N1-REGION', 'N2-REGION', '5\'J-REGION', '3\'V-REGION', 'N-REGION', 'D-REGION')


calculate_all_lengths <- function(data, regions) {
  do.call(rbind, lapply(regions, calculate_mean_length, data = data))
}

mean_lengths_list <- lapply(data_list, calculate_all_lengths, regions = region_names)


all_mean_lengths <- do.call(rbind, mean_lengths_list)

all_mean_lengths$Dataset <- rep(names(mean_lengths_list), each = length(region_names))


print(all_mean_lengths)
library(reshape2)


wide_all_mean_lengths <- dcast(all_mean_lengths, Region ~ Dataset, value.var = "MeanLength")

# 输出结果
print(wide_all_mean_lengths )

write.csv(wide_all_mean_lengths ,"Output/Tables/AllTCRBCR_CDR3_split_lengths.csv")
```


```{r}
IGLmean_lengths <- rbind(
  calculate_mean_length(IGL_data, 'N1-REGION'),
  calculate_mean_length(IGL_data, 'N2-REGION'),
  calculate_mean_length(IGL_data, '5\'J-REGION'),
  calculate_mean_length(IGL_data, '3\'V-REGION'),
  calculate_mean_length(IGL_data, 'N-REGION'),
  calculate_mean_length(IGL_data, 'D-REGION')
)
IGLmean_lengths 
```

```{r}

```


```{r}
IGH_data
table(IGH_data$`D-GENE and allele`)
```


```{r}

# IGH_data$`D-GENE and allele`
IGH_cleandata=IGH_data

IGHDcleaned_strings <- gsub("^Homsap\\s+|\\*.*$", "", IGH_data$`D-GENE and allele`)
IGHJcleaned_strings <- gsub("^Homsap\\s+|\\*.*$", "", IGH_data$`J-GENE and allele`)

IGH_cleandata$Clean_IGHD<-IGHDcleaned_strings
IGH_cleandata$Clean_IGHJ<-IGHJcleaned_strings
print(cleaned_strings)
IGHx1none<-subset(IGH_cleandata,Clean_IGHD!="")
```

```{r}

```

```{r}
# table(IGH_data$)
```

### VDJcombined
```{r}
library(dplyr)
library(tidyverse)
library(reshape2)
subVDJ$IR_VDH<-paste0(subVDJ$IR_VDJ_1_v_call,subVDJ$IR_VDJ_1_d_call,subVDJ$IR_VDJ_1_j_call)

```

#----------------
### DJ combined

```{r}
library(dplyr)
library(tidyverse)
library(reshape2)
subVDJ$IR_D<-subVDJ$IR_VDJ_1_d_call
subVDJ$IR_D<-ifelse(subVDJ$IR_D=="","None",subVDJ$IR_D)
# table(subVDJ$IR_VDJ_1_d_call)
subVDJ<-subset(subVDJ,IR_D!="None")
subVDJx1<-subset(subVDJ,IR_D="None")
table(subVDJx1$AdjustedID)
subVDJx1none<-subset(BCRobs,IR_VDJ_1_d_call=="")
```

```{r}
subVDJx1gene<-subset(BCRobs,IR_VDJ_1_d_call!="")
```

```{r}
table(subVDJx1none$New_Body,subVDJx1none$Last_cell_type)
```


```{r}
table(subVDJx1none$IR_VDJ_1_d_call)
```
```{r}
table(subVDJx1none$AdjustedID)
```

```{r}
table(subVDJ$IR_D)
grep("J",subVDJ$IR_D)
```
```{r}
NewBCR=merge(IGHx1none[,c(2,87,88)],BCRobs,by.x="Sequence ID",by.y="Cellname")
```

```{r}
library(dplyr)
library(tidyr)
library(reshape2)

IGHDconvert_to_wide <- function(data, id_column, d_gene_column, j_gene_column, target_gene_column, gene_list) {
  data$BCRHDJ <- paste0(data[[d_gene_column]], "_", data[[j_gene_column]])
  subD_data <- data[grep("B", data[[id_column]]),]
  BCRHDJ <- table(subD_data$BCRHDJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRHDJ$Per <- BCRHDJ$Freq / sum(BCRHDJ$Freq) * 100
  separated_dataD <- BCRHDJ %>%
    separate(Var1, into = c("Dgene", "Jgene"), sep = "_")
  Dwide_data <- separated_dataD %>% select(-Freq) %>%
    spread(key = Dgene, value = Per)
  rownames(Dwide_data) <- Dwide_data$Jgene
  Dwide_data <- Dwide_data[, -1]
  Dwide_data[is.na(Dwide_data)] <- 0
  missing_cols <- setdiff(gene_list, colnames(Dwide_data))
  for (col in missing_cols) {
    Dwide_data[[col]] <- 0
  }
  Dwide_data <- Dwide_data[, gene_list]
  BCRHDFreq_exp <- rownames_to_column(Dwide_data, var = "Jgene")
  BCRHDJlongexp <- melt(BCRHDFreq_exp, id.vars = "Jgene", variable.name = "Dgene", value.name = "Per")
  BCRHDJlongexp$Sample <- paste0(BCRHDJlongexp$Dgene, "_", BCRHDJlongexp$Jgene)
  return(BCRHDJlongexp)
}

# Use the function with your specific data frame and gene list variables
IGHDJresult <-IGHDconvert_to_wide(NewBCR, "AdjustedID", "Clean_IGHD", "Clean_IGHJ", "IGH",IGHDIMGT$IGH[-27])


```

```{r}
 library(dplyr)

create_sample_bcrhdj <- function(input_df, id_col, bcrhdj_col, group_prefix) {
  # Select only rows where the ID column starts with the specified prefix
  sub_df <- input_df[grep(group_prefix, input_df[[id_col]]), ]

  # Create a table of frequencies for each ID and BCRHDJ combination
  sample_bcrhdj <- table(sub_df[[id_col]], sub_df[[bcrhdj_col]]) %>%
    as.data.frame() %>%
    arrange(desc(Freq))

  # Rename columns to a standard format
  colnames(sample_bcrhdj) <- c("AdjustedID", "VDJ", "Freq")

  # Calculate total frequencies grouped by ID
  total_freq_by_id <- sample_bcrhdj %>%
    group_by(AdjustedID) %>%
    summarise(total_freq = sum(Freq)) %>%
    ungroup()  # Ensure that the data frame is no longer grouped

  # Join the total frequencies back to the original data
  sample_bcrhdj <- left_join(sample_bcrhdj, total_freq_by_id, by = "AdjustedID")

  return(sample_bcrhdj)
}

NewBCR$BCRHDJ<-paste0(NewBCR$Clean_IGHD,"_",NewBCR$Clean_IGHJ)
IGHDJresult_df <- create_sample_bcrhdj(NewBCR, "AdjustedID", "BCRHDJ", "B")
print(IGHDJresult_df )
```


```{r}
library(dplyr)
library(tidyr)

calculate_correlations <- function(df, id_col, freq_col, total_freq_col, group_col) {
  # Extracting the week information from the ID column
  df$Week <- as.numeric(substring(df[[id_col]], 2, 5))
  
  # Calculating the percentage
  df$Per <- df[[freq_col]] / df[[total_freq_col]]
  
  # Performing correlation tests grouped by the specified column
  cor_results <- df %>%
    group_by_at(vars(one_of(group_col))) %>%
    summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
              p.value = cor.test(Week, Per, method = "spearman")$p.value) %>%
    ungroup() %>%
    mutate(p.adjusted = p.adjust(p.value, method = "BH")) %>%
    arrange(cor)
  
  # Filtering the data based on the adjusted p-value
  significant_results <- cor_results %>%
    filter(p.adjusted < 0.1) %>%
    arrange(cor)
  
  # Adding a column to indicate the direction of the correlation
  cor_results$VDJcor <- ifelse(cor_results$p.adjusted > 0.05, "NA",
                                ifelse(cor_results$cor > 0, "+", "-"))
  
  # Return the results
  list(cor_results = cor_results, significant_results = significant_results)
}

# To use the function, you'd call it with the appropriate column names for your dataframe:
IGHDJresult_dfcor<- calculate_correlations(IGHDJresult_df, "AdjustedID", "Freq", "total_freq", "VDJ")
print(IGHDJresult_dfcor$cor_results)
print(IGHDJresult_dfcor$significant_results)
```
```{r}
library(dplyr)
library(tidyr)

process_cor_results <- function(cor_results_df, cor_col, group_col) {
  # Assign size factors based on the absolute value of the correlation
  cor_results_df$size_factor <- cut(
    abs(cor_results_df[[cor_col]]),
    breaks = c(-Inf, 0.5, 0.6, 0.7, 0.8, Inf),
    labels = c("0-0.5", "0.5-0.6", "0.6-0.7", "0.7-0.8", "0.8-1.0")
  )

  # Create a copy of the VDJ column for later use
  cor_results_df <- cor_results_df %>%
    mutate(VDJ2 = .data[[group_col]]) %>%
    separate(.data[[group_col]], into = c("Dgene", "Jgene"), sep = "_") %>%
    mutate(VDJ = VDJ2) %>%
    mutate(VDJ2 = paste0(Jgene, "_", Dgene))

  # Filter out rows with "NA" in the correlation significance column
  filtered_cor_results <- subset(cor_results_df, VDJcor != "NA")

  # Remove duplicates based on the VDJ2 column
  unique_cor_results <- filtered_cor_results[!duplicated(filtered_cor_results$VDJ2),]

  return(unique_cor_results)
}

SIGHDJresult_dfcorsig<- process_cor_results(IGHDJresult_dfcor$cor_results, "cor", "VDJ")
# print(final_results)
```

```{r}
P1DJ=ggplot() +
  geom_tile(data = IGHDJresult , aes(x = Dgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRH", x = "D gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))+  geom_text(data = SIGHDJresult_dfcorsig, 
            aes(x = Dgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4,"0.5-0.6"=5,  "0.6-0.7" = 6, "0.7-0.8"= 7,"0.8-0.9"= 8,"0.8-1.0"=9))
P1DJ
```
```{r}
ggsave(plot = P1DJ,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Main_Figure5/Figure5H_BCRH_all_DJcombined.pdf",width = 8,height = 4)
ggsave(plot = P1DJ,"Output/Plot/Figure5H_BCRH_all_DJcombined.pdf",width = 8,height = 4)
```

```{r}
subVDJ$BCRHDJ<-paste0(subVDJ$IR_D,"_",subVDJ$IR_VDJ_1_j_call)
subD_data =subVDJ[grep("B", subVDJ[["AdjustedID"]]),]
BCRHDJ = table(subD_data$BCRHDJ) %>% as.data.frame() %>% arrange(desc(Freq))
BCRHDJ$Per = BCRHDJ$Freq / sum(BCRHDJ$Freq)*100

separated_dataD <- BCRHDJ %>%
  separate(Var1, into = c("Dgene", "Jgene"), sep = "_")

Dwide_data <- separated_dataD %>% select(-Freq) %>%
  spread(key = Dgene, value = Per)

rownames(Dwide_data) <- Dwide_data$Jgene
Dwide_data <- Dwide_data[, -1]
Dwide_data[is.na(Dwide_data)] <- 0
missing_cols <- setdiff(IGHDIMGT$IGH[-27], colnames(Dwide_data))

  for(col in missing_cols) {
    Dwide_data[[col]] <- 0
  }

  Dwide_data <- Dwide_data[,IGHDIMGT$IGH[-27]]
BCRHDFreq_exp=rownames_to_column(Dwide_data)
BCRHDJlongexp <- melt(BCRHDFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")

colnames(BCRHDJlongexp)<-c("Jgene","Dgene","Per")
BCRHDJlongexp$Sample=paste0(BCRHDJlongexp$Dgene,"_",BCRHDJlongexp$Jgene)
```





```{r}
subBCRHh1=subVDJ[grep("B",subVDJ$AdjustedID),]
Sample_BCRHDJ=table(subBCRHh1$AdjustedID,subBCRHh1$BCRHDJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_BCRHDJ)<-c("AdjustedID","VDJ","Freq")

DJtotal_freq_by_id <- Sample_BCRHDJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_BCRHDJ <- left_join(Sample_BCRHDJ, DJtotal_freq_by_id, by = "AdjustedID")
```

```{r}
Sample_BCRHDJ$Week<-substring(Sample_BCRHDJ$AdjustedID,2,5) %>% as.numeric()
Sample_BCRHDJ$Per=Sample_BCRHDJ$Freq/Sample_BCRHDJ$total_freq
Sample_BCRHDJ_cor_results <- Sample_BCRHDJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
Sample_BCRHDJ_cor_results<-Sample_BCRHDJ_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(Sample_BCRHDJ_cor_results)
subset(Sample_BCRHDJ_cor_results,p.adjusted<0.1) %>% arrange(cor)
Sample_BCRHDJ_cor_results$VDJcor<-ifelse(Sample_BCRHDJ_cor_results$p.adjusted>0.1,"NA",ifelse(Sample_BCRHDJ_cor_results$cor>0,"+","-"))
```



```{r}
 Sample_BCRHVJ 
```

```{r}
Sample_BCRHDJ_cor_results
```


```{r}
# 
#   return(wide_data)
# }

Sample_BCRHDJ_cor_results$size_factor <- cut(abs(Sample_BCRHDJ_cor_results$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))

Sample_BCRHDJ_cor_results=Sample_BCRHDJ_cor_results%>%mutate(VDJ2=VDJ) %>% 
  separate(VDJ, into = c("Dgene", "Jgene"), sep = "_")%>%mutate(VDJ=VDJ2) %>% 
mutate(VDJ2=paste0(Jgene,"_",Dgene))
Sample_BCRHDJ_cor_results2<-subset(Sample_BCRHDJ_cor_results,VDJcor!="NA")
Sample_BCRHDJ_cor_results2<-Sample_BCRHDJ_cor_results2[!duplicated(Sample_BCRHDJ_cor_results2$VDJ2),]
```


```{r}
ggplot() +
  geom_tile(data = BCRHDJlongexp, aes(x = Dgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRH", x = "D gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))+  geom_text(data = Sample_BCRHDJ_cor_results2, 
            aes(x = Dgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4,"0.5-0.6"=5,  "0.6-0.7" = 6, "0.7-0.8"= 7,"0.8-0.9"= 8,"0.8-1.0"=9))


```

#----------------

####  BCRHlongexp
```{r}
BCRHFreq_exp=VDJfun(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)


BCRHFreq_exp=rownames_to_column(BCRHFreq_exp)
BCRHlongexp <- melt(BCRHFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")
colnames(BCRHlongexp)<-c("Jgene","Vgene","Per")
BCRHlongexp$Sample=paste0(BCRHlongexp$Vgene,"_",BCRHlongexp$Jgene)
BCRHlongexp$VDJsig<-BCRHlongexp$Sample %in% as.character(top_VDJ)

BCRHlongexp$VDJsig <- ifelse(BCRHlongexp$VDJsig, BCRHlongexp$Sample, "Others")
```



```{r}
library(ggplot2)


PBMCBCRH_cor_results2$size_factor <- cut(abs(PBMCBCRH_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


BCRHp1 <- ggplot() +
  geom_tile(data = BCRHlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRH", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCBCRH_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(BCRHp1)

ggsave(plot = BCRHp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Main_Figure5/Figure5H_BCRH_all_VJcombined.pdf",width = 8,height = 4)
```



#### PBMCBCRH_cor_results
```{r}
dim(PBMCBCRH_cor_results)
PBMCBCRH_cor_results=PBMCBCRH_cor_results %>% separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")
PBMCBCRH_cor_results2=subset(PBMCBCRH_cor_results,p.value<0.05&abs(cor)>0.6)
dim(BCRHlongexp)
# setdiff(BCRHlongexp$Sample,PBMCBCRH_cor_results$VDJ)
```


#### Plot:cor percentage heatmmap 
```{r}
# BCRHlongexp$Jgene<-factor(BCRHlongexp$Jgene,levels = rev($IMGT_gene_name))
```
```{r}
PBMCBCRH_cor_results2
```


```{r}
library(ggplot2)


PBMCBCRH_cor_results2$size_factor <- cut(abs(PBMCBCRH_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


BCRHp1 <- ggplot() +
  geom_tile(data = BCRHlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRH", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCBCRH_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(BCRHp1)

ggsave(plot = BCRHp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Main_Figure5/Figure5H_BCRH_all_VJcombined.pdf",width = 8,height = 4)
```
```{r}

```


#### Plot:test ggplot cluster 
```{r}
# library(reshape2)
# mat_data <- dcast(long_data, Vgene ~ Jgene, value.var = "Per")
# rownames(mat_data) <- mat_data$Vgene
# mat_data <- mat_data[, -1]
# mat_data <- as.matrix(mat_data)
# clusters_row <- hclust(dist(mat_data), method = "ave")
# clusters_col <- hclust(dist(t(mat_data)), method = "ave")
# long_data_clustered <- melt(mat_data)
# colnames(long_data_clustered) <- c("Vgene", "Jgene", "Per")
```



```{r}
# long_data_clustered$Vgene<-as.character(long_data_clustered$Vgene)
# long_data_clustered$Jgene<-as.character(long_data_clustered$Jgene)
```



```{r}
# P1 <- ggplot(long_data_clustered, aes(x =Jgene, y =  Vgene, fill = Per)) +
#   geom_tile(color = "white") +
#   scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
#                        na.value = "white") +
#   theme_minimal() +
#   labs(title = "BCRH", x = "V gene", y = "J gene") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1),
#         plot.title = element_text(hjust = 0.5)) +
#   scale_x_dendrogram(labels = T, hclust = clusters_col) +
#   scale_y_dendrogram(labels = T, hclust = clusters_row) 
# 
# print(P1)
```


```{r}
# ggplot()+geom_tile(data=PBMCBCRHVJ_separated,aes(Vgene,Jgene,fill=0),color="gray52",fill="white",size=1)+geom_tile(data=PBMCBCRHVJ_separated ,aes(Vgene,Jgene,value=Per,fill=Per))+ theme(axis.text.x=element_text(angle=90, hjust=1)) + scale_fill_gradientn(colours =rev(colorRampPalette(c('#C6403D', 'white','#246BAE'))(111)))+theme(axis.text.y=element_text(size = 12,color=c(rep("#246BAE",4),rep("#D0AB39",4),rep("#3FBDA7",4),rep("#C6403D",4))))+
#   guides(fill=guide_colorbar(barheight = 28))+theme_bw()+
#   theme(panel.grid = element_blank(),
#         panel.border = element_blank(),
#         axis.text.x = element_text(angle = 60,
#                                    hjust=1,vjust = 1))
```

#----Per Sample BCRK----

### Get Sample percentage

```{r}
subVDJ$BCRKVJ<-paste0(subVDJ$IR_VJ_1_v_call,"_",subVDJ$IR_VJ_1_j_call)
```

```{r}
subVDJ[grep("B",subVDJ$AdjustedID),] %>% .[grep("IGKV",.$BCRKVJ),]
```

```{r}
subBCRKh1=subVDJ[grep("B",subVDJ$AdjustedID),] %>% .[grep("IGKV",.$BCRKVJ),]
Sample_BCRKVJ=table(subBCRKh1$AdjustedID,subBCRKh1$BCRKVJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_BCRKVJ)<-c("AdjustedID","VDJ","Freq")

BCRKVJ_total_freq_by_id <- Sample_BCRKVJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_BCRKVJ <- left_join(Sample_BCRKVJ, BCRKVJ_total_freq_by_id, by = "AdjustedID")


Sample_BCRKVJ <- Sample_BCRKVJ %>% 
  mutate(Per = Freq / total_freq*100)

Sample_BCRKVJ <- Sample_BCRKVJ %>% arrange(AdjustedID)
Sample_BCRKVJ$Week=substring(Sample_BCRKVJ$AdjustedID,2,5) %>% as.numeric()
length(unique(Sample_BCRKVJ$VDJ))
```
```{r}
library(dplyr)
library(ggplot2)

Sample_BCRKVJ$Week <- as.numeric(Sample_BCRKVJ$Week)
Sample_BCRKVJ$Per <- as.numeric(Sample_BCRKVJ$Per)


topA_VDJ <- Sample_BCRKVJ %>% 
  group_by(VDJ) %>%
  summarise(Total_Per = sum(Per)) %>%
  top_n(10, Total_Per) %>%
  pull(VDJ)

Sample_BCRKVJ <- Sample_BCRKVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% topA_VDJ, VDJ, "Other"))
Sample_BCRKVJ$VDJ_Top10<-factor(Sample_BCRKVJ$VDJ_Top10,levels = c(as.character(topA_VDJ),"Other"))
```

```{r}
ggplot(Sample_BCRKVJ, aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  geom_line() +
  scale_color_manual(values = c(rainbow(10), "gray")) +
  labs(x = "Week", y = "Per", color = "VDJ") +
  theme_classic()
```
#### corBCRK
```{r}
PBMCBCRK_cor_results <- Sample_BCRKVJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
PBMCBCRK_cor_results<-PBMCBCRK_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(PBMCBCRK_cor_results)
subset(PBMCBCRK_cor_results,p.value<0.05) %>% arrange(cor)
PBMCBCRK_cor_results$VDJcor<-ifelse(PBMCBCRK_cor_results$p.value>0.05,"NA",ifelse(PBMCBCRK_cor_results$cor>0,"+","-"))
```

### BCRKfunction Data

```{r}
library(ggplot2)
library(reshape2) # for melt function

BCRKVDJfun <- function(data,pattern, BCRKinIMGT=BCRKIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRKVJ = table(sub_data$BCRKVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRKVJ$Per = BCRKVJ$Freq / sum(BCRKVJ$Freq)*100
  
  separated_data <- BCRKVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(BCRKinIMGT$IMGT_gene_name, colnames(wide_data))
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  
  wide_data <- wide_data[,BCRKinIMGT$IMGT_gene_name]
  
  return(wide_data)
}
```

```{r}
BCRKIMGTsub=IGKVIMGT[IGKVIMGT$IMGT.gene.name %in% subVDJ$IR_VJ_1_v_call,] %>% arrange(IMGT.gene.order)
colnames(BCRKIMGTsub)<-c("IMGT_gene_name","IMGT.gene.order")
```
####  BCRKlongexp
```{r}
subKVDJ=subVDJ%>% .[grep("IGKV",.$BCRKVJ),]
BCRKFreq_exp=BCRKVDJfun(data = subKVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)


BCRKFreq_exp=rownames_to_column(BCRKFreq_exp)
BCRKlongexp <- melt(BCRKFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")
colnames(BCRKlongexp)<-c("Jgene","Vgene","Per")
BCRKlongexp$Sample=paste0(BCRKlongexp$Vgene,"_",BCRKlongexp$Jgene)
BCRKlongexp$VDJsig<-BCRKlongexp$Sample %in% as.character(top_VDJ)


BCRKlongexp$VDJsig <- ifelse(BCRKlongexp$VDJsig, BCRKlongexp$Sample, "Others")
```

#### PBMCBCRK_cor_results
```{r}
dim(PBMCBCRK_cor_results)
PBMCBCRK_cor_results=PBMCBCRK_cor_results %>% separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")
PBMCBCRK_cor_results2=subset(PBMCBCRK_cor_results,p.value<0.05&abs(cor)>0.6)
dim(BCRKlongexp)
# setdiff(BCRKlongexp$Sample,PBMCBCRK_cor_results$VDJ)
```
```{r}
IGKVIMGT$IMGT.gene.name
```


#### Plot:cor percentage heatmmap 
```{r}
BCRKlongexp$Jgene<-factor(BCRKlongexp$Jgene,levels = rev(IGKJIMGT$IMGT_gene_name))
```
```{r}
PBMCBCRK_cor_results2
```

```{r}
library(ggplot2)


PBMCBCRK_cor_results2$size_factor <- cut(abs(PBMCBCRK_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


BCRKp1 <- ggplot() +
  geom_tile(data = BCRKlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRK", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCBCRK_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c( "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(BCRKp1)

ggsave(plot = BCRKp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Main_Figure5/Figure5H_BCRK_all_VJcombined.pdf",width = 8,height = 4)
```

```{r}
BCRKVDJheatmap <- function(data,pattern, BCRKinIMGT=BCRKIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRKVJ = table(sub_data$BCRKVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRKVJ$Per = BCRKVJ$Freq / sum(BCRKVJ$Freq)*100
  
  separated_data <- BCRKVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(BCRKinIMGT$IMGT_gene_name, colnames(wide_data))
  
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  wide_data<-wide_data[,BCRKinIMGT$IMGT_gene_name]
  pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c('#C6403D', "#f6c619","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  return(wide_data)
}
```

```{r}
BCRKVDJheatmap(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)
```

# ----BCRH----

```{r}

subBCRHvdj=subBCRHobs2[, c("AdjustedID", "New_Body", colnames(subBCRHobs2)[grep("VDJ", colnames(subBCRHobs2))])]
subBCRHvdj$BCRHvj=paste0(subBCRHvdj$IR_VDJ_1_v_call,"_",subBCRHvdj$IR_VDJ_1_j_call)
subBCRHvdj$aalength=nchar(subBCRHvdj$IR_VDJ_1_junction_aa)
subBCRHvdjlen=table(subBCRHvdj$BCRHvj,subBCRHvdj$aalength) %>% as.data.frame()
```

```{r}
# subset(subBCRHvdjlen,Len==17&Freq>median((subset(subBCRHvdjlen,Len==17&Freq>0)$Freq)))
# max(subset(subBCRHvdjlen,Len==17&Freq>0)$Freq)
```
```{r}
PBMCBCRH_cor_results2$VJgene=paste0(PBMCBCRH_cor_results2$Vgene,"_",PBMCBCRH_cor_results2$Jgene)
# intersect(PBMCBCRH_cor_results2$VJgene,subset(subBCRHvdjlen,Len==17&Freq>median((subset(subBCRHvdjlen,Len==17&Freq>0)$Freq)))$VDJ)
```


```{r}
colnames(subBCRHvdjlen)<-c("VDJ","Len","Freq")
subBCRHvdjlen$Percentage=subBCRHvdjlen$Freq/sum(subBCRHvdjlen$Freq)*100

P2=ggplot(subBCRHvdjlen)+geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRH", x = "V gene", y = "J gene",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
P2
```

```{r}
subBCRHvdjlen_JV=subBCRHvdjlen%>%mutate(VDJ2=VDJ) %>% 
  separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")%>%mutate(VDJ=VDJ2) %>% 
mutate(VDJ2=paste0(Jgene,"_",Vgene)) 

subBCRHvdjlen_JV[subBCRHvdjlen_JV$Jgene %in% c("IGHJ6","IGHJ3","IGHJ4"),] %>% 
ggplot()+geom_tile(aes(Len,VDJ2,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRH", x = "CDR3 length", y = "J-V pair",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) ->PBCH_len_J
PBCH_len_J
ggsave(plot = PBCH_len_J,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Sup_Figure5/S30/S30_BCRH_IGVJcombined.pdf",height = 14,width = 8)
```
```{r}
#ggsave(plot=PBCH_len_J,"Figure/PBCH_len_J.pdf",width = 6,height = 15)
```



```{r}
subBCRHvdjlen[grep("IGHJ6|IGHJ4",subBCRHvdjlen$VDJ),]
```

```{r}
subBCRHvdjlen[grep("IGHJ6",subBCRHvdjlen$VDJ),] %>%
  ggplot() +geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRH", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}

```

```{r}

subBCRHvdjlen[subBCRHvdjlen$VDJ %in% intersect(PBMCBCRH_cor_results2$VJgene,subset(subBCRHvdjlen,Len==17&Freq>median((subset(subBCRHvdjlen,Len==17&Freq>0)$Freq)))$VDJ),] %>%
  ggplot() +geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRH", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}
# subBCRHvdjlen=table(subBCRHvdj$BCRHvj,subBCRHvdj$aalength) %>% as.data.frame()
```


```{r}
PBMCBCRH_cor_results2$VDJgene=paste0(PBMCBCRH_cor_results2$Vgene,"_",PBMCBCRH_cor_results2$Jgene)
BCRHlongexpsig=merge(BCRHlongexp,PBMCBCRH_cor_results2,by.x="Sample",by.y="VDJgene")
subset(BCRHlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10)
# max_five <- sort(BCRHlongexp$Per, decreasing = TRUE) %>% head( 10)

subset(BCRHlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subset(BCRHlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subBCRHsortgene=subset(BCRHlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor) %>% .$Sample
subBCRHvdjlen[subBCRHvdjlen$VDJ %in%subBCRHsortgene ,] %>% mutate(VDJ=factor(VDJ,levels=rev(subBCRHsortgene)))%>%
  ggplot() +geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                                                                                        na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRH", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}
# subBCRHvdjlenmax
```

### Plot:BCRH All len
```{r}
subBCRHvdjlenmax <- subBCRHvdjlen %>%
  group_by(VDJ) %>%
  mutate(Percentage_max = max(Percentage)) %>%
  filter(Percentage == Percentage_max)
m1=merge(subBCRHvdjlenmax,BCRHlongexpsig,by.x="VDJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VDJ %in% subBCRHsortgene,subBCRHsortgene,"Others")) 
m1<-m1[!duplicated(m1$VDJ),]
length(unique(m1$VDJ))
m1$VDJcol <- subBCRHsortgene[match(m1$VDJ, subBCRHsortgene)]
m1$VDJcol[is.na(m1$VDJcol)] <- "Others BCRH V-J"
table(m1$VDJcol)
setdiff( subBCRHsortgene,m1$VDJ )

library(ggsci)

P3BCRHplot=m1%>% mutate(Rcor=ifelse(cor<0,"R<0","R>0")) %>% mutate(VDJcol=factor(VDJcol,levels=c(subBCRHsortgene,"Others BCRH V-J"))) %>% 
  ggplot() +
  geom_point(aes(Len,cor,size=Per,color=VDJcol,shape=Rcor)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC BCRH", x = "BCRH V-J CDR3 max len", y = " Percentage in All(%)",size="Percentage in BCRH V-J(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),aspect.ratio = 0.6)+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))+scale_color_futurama()
P3BCRHplot
```
```{r}
# #ggsave(plot=P3BCRHplot,"Figure/P3BCRHplot.pdf",width = 8,height = 6)
```
### Plot:BCRH All len
```{r}
merge(subBCRHvdjlen,BCRHlongexpsig,by.x="VDJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VDJ %in% subBCRHsortgene,subBCRHsortgene,"Others")) %>% mutate(Percentage2=ifelse(cor <0,-Percentage,Percentage)) %>% 
  ggplot() +
  geom_point(aes(Len,Percentage2,color=VDJcol)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC BCRH", x = "VDJ CDR3 max len", y = "Percentage(%)",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))
```


#----BCRK----
```{r}
subBCRKvdj=subBCRHobs2[, c("AdjustedID", "New_Body", colnames(subBCRHobs2)[grep("VJ", colnames(subBCRHobs2))])]
subBCRKvdj$BCRKvj=paste0(subBCRKvdj$IR_VJ_1_v_call,"_",subBCRKvdj$IR_VJ_1_j_call)
subBCRKvdj$aalength=nchar(subBCRKvdj$IR_VJ_1_junction_aa)
subBCRKvdjlen=table(subBCRKvdj$BCRKvj,subBCRKvdj$aalength) %>% as.data.frame()
max(subBCRKvdjlen$Percentage)
colnames(subBCRKvdjlen)<-c("VJ","Len","Freq")
subBCRKvdjlen$Percentage=subBCRKvdjlen$Freq/sum(subBCRKvdjlen$Freq)*100
```
### Plot:BCRK All len
```{r}
subBCRKvdjlen_JV=subBCRKvdjlen%>%mutate(VJ2=VJ) %>% 
  separate(VJ, into = c("Vgene", "Jgene"), sep = "_")%>%mutate(VJ=VJ2) %>% 
  mutate(VJ2=paste0(Jgene,"_",Vgene)) 
PlenBCRK=ggplot(subBCRKvdjlen_JV)+geom_tile(aes(Len,VJ2,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619", '#246BAE')), 
                                                                                                     na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRK", x = "CDR3 Length", y = "J gene",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        )
PlenBCRK
```

```{r}
#ggsave(plot=PlenBCRK,"Figure/PlenBCRK.pdf",width = 6,height = 24)
```

```{r}
# #ggsave(plot=PlenBCRK,"Figure/PlenBCRK.pdf")
```

```{r}
PBMCBCRK_cor_results2$VJgene[1:10]
```
```{r}
PBMCBCRK_cor_results2$VJgene=paste0(PBMCBCRK_cor_results2$Vgene,"_",PBMCBCRK_cor_results2$Jgene)
BCRKlongexpsig=merge(BCRKlongexp,PBMCBCRK_cor_results2,by.x="Sample",by.y="VJgene")
subset(BCRKlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10)
# max_five <- sort(BCRKlongexp$Per, decreasing = TRUE) %>% head( 10)

subset(BCRKlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subset(BCRKlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```
### Plot:BCRK ten len
```{r}
subBCRKsortgene=subset(BCRKlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor) %>% .$Sample
subBCRKvdjlen[subBCRKvdjlen$VJ %in%subBCRKsortgene ,] %>% mutate(VJ=factor(VJ,levels=rev(subBCRKsortgene)))%>%
  ggplot() +geom_tile(aes(Len,VJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                                                                                         na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRK", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}
subBCRKvdjlen
```

```{r}
subBCRKvdjlenmax <- subBCRKvdjlen %>%
  group_by(VJ) %>%
  mutate(Percentage_max = max(Percentage)) %>%
  filter(Percentage == Percentage_max)
m1=merge(subBCRKvdjlenmax,BCRKlongexpsig,by.x="VJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VJ %in% subBCRKsortgene,subBCRKsortgene,"Others")) 
m1<-m1[!duplicated(m1$VJ),]
length(unique(m1$VJ))
m1$VDJcol <- subBCRKsortgene[match(m1$VJ, subBCRKsortgene)]
m1$VDJcol[is.na(m1$VDJcol)] <- "Others BCRK V-J"
table(m1$VDJcol)
setdiff( subBCRKsortgene,m1$VJ )

library(ggsci)

P3BCRKplot=m1%>% mutate(Rcor=ifelse(cor<0,"R<0","R>0")) %>% mutate(VDJcol=factor(VDJcol,levels=c(subBCRKsortgene,"Others BCRK V-J"))) %>% 
  ggplot() +
  geom_point(aes(Len,cor,size=Per,color=VDJcol,shape=Rcor)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC BCRK", x = "BCRK V-J CDR3 max len", y = " Percentage in All BCRK V-J(%)",size="Percentage in BCRK V-J(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),aspect.ratio = 0.6)+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))+scale_color_futurama()
P3BCRKplot
```

```{r}
#ggsave(plot=P3BCRKplot,"Figure/P3BCRKplot.pdf")
```

#----BCRL----

### Get Sample percentage

```{r}
subVDJ$BCRLVJ<-paste0(subVDJ$IR_VJ_1_v_call,"_",subVDJ$IR_VJ_1_j_call)
```

```{r}
subVDJ[grep("B",subVDJ$AdjustedID),] %>% .[grep("IGLV",.$BCRLVJ),]
```

```{r}
subBCRLh1=subVDJ[grep("B",subVDJ$AdjustedID),] %>% .[grep("IGLV",.$BCRLVJ),]
Sample_BCRLVJ=table(subBCRLh1$AdjustedID,subBCRLh1$BCRLVJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_BCRLVJ)<-c("AdjustedID","VDJ","Freq")

BCRLVJ_total_freq_by_id <- Sample_BCRLVJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_BCRLVJ <- left_join(Sample_BCRLVJ, BCRLVJ_total_freq_by_id, by = "AdjustedID")


Sample_BCRLVJ <- Sample_BCRLVJ %>% 
  mutate(Per = Freq / total_freq*100)

Sample_BCRLVJ <- Sample_BCRLVJ %>% arrange(AdjustedID)
Sample_BCRLVJ$Week=substring(Sample_BCRLVJ$AdjustedID,2,5) %>% as.numeric()
length(unique(Sample_BCRLVJ$VDJ))
unique(Sample_BCRLVJ$AdjustedID)
```
```{r}
library(dplyr)
library(ggplot2)

Sample_BCRLVJ$Week <- as.numeric(Sample_BCRLVJ$Week)
Sample_BCRLVJ$Per <- as.numeric(Sample_BCRLVJ$Per)


topA_VDJ <- Sample_BCRLVJ %>% 
  group_by(VDJ) %>%
  summarise(Total_Per = sum(Per)) %>%
  top_n(10, Total_Per) %>%
  pull(VDJ)

Sample_BCRLVJ <- Sample_BCRLVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% topA_VDJ, VDJ, "Other"))
Sample_BCRLVJ$VDJ_Top10<-factor(Sample_BCRLVJ$VDJ_Top10,levels = c(as.character(topA_VDJ),"Other"))
```

```{r}
ggplot(Sample_BCRLVJ, aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  geom_line() +
  scale_color_manual(values = c(rainbow(10), "gray")) +
  labs(x = "Week", y = "Per", color = "VDJ") +
  theme_classic()
```

#### corBCRL
```{r}
PBMCBCRL_cor_results <- Sample_BCRLVJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
PBMCBCRL_cor_results<-PBMCBCRL_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(PBMCBCRL_cor_results)
subset(PBMCBCRL_cor_results,p.value<0.05) %>% arrange(cor)
PBMCBCRL_cor_results$VDJcor<-ifelse(PBMCBCRL_cor_results$p.value>0.05,"NA",ifelse(PBMCBCRL_cor_results$cor>0,"+","-"))
```
### BCRLfunction Data

```{r}
library(ggplot2)
library(reshape2) # for melt function

BCRLVDJfun <- function(data,pattern, BCRLinIMGT=BCRLIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRLVJ = table(sub_data$BCRLVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRLVJ$Per = BCRLVJ$Freq / sum(BCRLVJ$Freq)*100
  
  separated_data <- BCRLVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(BCRLinIMGT$IMGT_gene_name, colnames(wide_data))
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  
  wide_data <- wide_data[,BCRLinIMGT$IMGT_gene_name]
  
  return(wide_data)
}
```

```{r}
BCRLIMGTsub=IGLVIMGT[IGLVIMGT$IMGT.gene.name %in% subVDJ$IR_VJ_1_v_call,] %>% arrange(IMGT.gene.order)
colnames(BCRLIMGTsub)<-c("IMGT_gene_name","IMGT.gene.order")
```
####  BCRLlongexp
```{r}
subLVDJ=subVDJ%>% .[grep("IGLV",.$BCRLVJ),]
BCRLFreq_exp=BCRLVDJfun(data = subLVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)


BCRLFreq_exp=rownames_to_column(BCRLFreq_exp)
BCRLlongexp <- melt(BCRLFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")
colnames(BCRLlongexp)<-c("Jgene","Vgene","Per")
BCRLlongexp$Sample=paste0(BCRLlongexp$Vgene,"_",BCRLlongexp$Jgene)
BCRLlongexp$VDJsig<-BCRLlongexp$Sample %in% as.character(top_VDJ)


BCRLlongexp$VDJsig <- ifelse(BCRLlongexp$VDJsig, BCRLlongexp$Sample, "Others")
```

#### PBMCBCRL_cor_results
```{r}
dim(PBMCBCRL_cor_results)
PBMCBCRL_cor_results=PBMCBCRL_cor_results %>% separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")
PBMCBCRL_cor_results2=subset(PBMCBCRL_cor_results,p.value<0.05&abs(cor)>0.6)
dim(BCRLlongexp)
# setdiff(BCRLlongexp$Sample,PBMCBCRL_cor_results$VDJ)
```
#### Plot:cor percentage heatmmap 
```{r}
BCRLlongexp$Jgene<-factor(BCRLlongexp$Jgene,levels = rev(IGLJIMGT$IMGT.gene.name))
```


```{r}
library(ggplot2)


PBMCBCRL_cor_results2$size_factor <- cut(abs(PBMCBCRL_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


BCRLp1 <- ggplot() +
  geom_tile(data = BCRLlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC All BCRL", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCBCRL_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(BCRLp1)

ggsave(plot = BCRLp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Main_Figure5/Figure5H_BCRL_all_VJcombined.pdf",width = 8,height = 4)
```


```{r}
BCRLVDJheatmap <- function(data,pattern, BCRLinIMGT=BCRLIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  BCRLVJ = table(sub_data$BCRLVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  BCRLVJ$Per = BCRLVJ$Freq / sum(BCRLVJ$Freq)*100
  
  separated_data <- BCRLVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(BCRLinIMGT$IMGT_gene_name, colnames(wide_data))
  
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  wide_data<-wide_data[,BCRLinIMGT$IMGT_gene_name]
  pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c('#C6403D', "#f6c619","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  return(wide_data)
}
```

```{r}
BCRLVDJheatmap(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)
```


#----BCRL----
```{r}
subBCRLvdj=subBCRHobs2[, c("AdjustedID", "New_Body", colnames(subBCRHobs2)[grep("VJ", colnames(subBCRHobs2))])]
subBCRLvdj$BCRLvj=paste0(subBCRLvdj$IR_VJ_1_v_call,"_",subBCRLvdj$IR_VJ_1_j_call)
subBCRLvdj$aalength=nchar(subBCRLvdj$IR_VJ_1_junction_aa)
subBCRLvdj=subBCRLvdj[grep("B",subBCRLvdj$AdjustedID),]
unique(subBCRLvdj$AdjustedID)
subBCRLvdj=subBCRLvdj[grep("IGLV",subBCRLvdj$BCRLvj),]
subBCRLvdjlen=table(subBCRLvdj$BCRLvj,subBCRLvdj$aalength) %>% as.data.frame()
# max(subBCRLvdjlen$Percentage)
colnames(subBCRLvdjlen)<-c("VJ","Len","Freq")
subBCRLvdjlen$Percentage=subBCRLvdjlen$Freq/sum(subBCRLvdjlen$Freq)*100
```

### Plot:BCRL All len
```{r}
subBCRLvdjlen
PlenBCRL=ggplot(subBCRLvdjlen)+geom_tile(aes(Len,VJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619", '#246BAE')), 
                                                                                                           na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRL", x = "CDR3 Length", y = "J gene",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank())
PlenBCRL
```
```{r}
PBMCBCRL_cor_results2$VJgene=paste0(PBMCBCRL_cor_results2$Vgene,"_",PBMCBCRL_cor_results2$Jgene)
BCRLlongexpsig=merge(BCRLlongexp,PBMCBCRL_cor_results2,by.x="Sample",by.y="VJgene")
subset(BCRLlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10)
# max_five <- sort(BCRLlongexp$Per, decreasing = TRUE) %>% head( 10)

subset(BCRLlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

### Plot:BCRL ten len
```{r}
subBCRLsortgene=subset(BCRLlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor) %>% .$Sample
subBCRLvdjlen[subBCRLvdjlen$VJ %in%subBCRLsortgene ,] %>% mutate(VJ=factor(VJ,levels=rev(subBCRLsortgene)))%>%
  ggplot() +geom_tile(aes(Len,VJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619","white", '#246BAE')), 
                                                                                        na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC BCRL", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```


```{r}
subBCRLvdjlenmax <- subBCRLvdjlen %>%
  group_by(VJ) %>%
  mutate(Percentage_max = max(Percentage)) %>%
  filter(Percentage == Percentage_max)
m1=merge(subBCRLvdjlenmax,BCRLlongexpsig,by.x="VJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VJ %in% subBCRLsortgene,subBCRLsortgene,"Others")) 
m1<-m1[!duplicated(m1$VJ),]
length(unique(m1$VJ))
m1$VDJcol <- subBCRLsortgene[match(m1$VJ, subBCRLsortgene)]
m1$VDJcol[is.na(m1$VDJcol)] <- "Others BCRL V-J"
table(m1$VDJcol)
setdiff( subBCRLsortgene,m1$VJ )

library(ggsci)

P3BCRLplot=m1%>% mutate(Rcor=ifelse(cor<0,"R<0","R>0")) %>% mutate(VDJcol=factor(VDJcol,levels=c(subBCRLsortgene,"Others BCRL V-J"))) %>% 
  ggplot() +
  geom_point(aes(Len,cor,size=Per,color=VDJcol,shape=Rcor)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC BCRL", x = "BCRL V-J CDR3 max len", y = " Rcor",size="Percentage in BCRL V-J(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),aspect.ratio = 0.6)+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))+scale_color_futurama()
P3BCRLplot
```


#---- Save and load data----

```{r}
# rm(list = ls())
### 2023-10-28 23:30:58
# load("/data1/maolp/Codeman/Project/Main_Gao_ScanpyProject20231003/Gao_plot/Figure3_TCRBCR_CDR3/DATA/BCR_MainBCR2_20231028.RData")
```