---
title: "CD4 TCR"
output: html_document
date: "2023-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

```{r}
# load("/data1/maolp/Codeman/Project/Main_Gao_ScanpyProject20231003/Gao_plot/Figure3_TCRBCR_CDR3/DATA/CD4_MainTCR2_20231028.RData")
```


#----R load----
### Load R packages
```{r}
library(tidyverse)
library(pheatmap)
```

#----Data load----


```{r}
TCRdir="TCR_all/"
TCRsamples=list.files(path = TCRdir,"csv")
AllTCR=data.frame()
for (line in TCRsamples) {
#   print(line)
  name=gsub("_TCR_filtered_contig_annotations.csv","",line)
  # print(name)
  if (!name %in% c("C27","Z28")){
     Test_TCR<-read.csv(file.path(TCRdir,line))
     Test_TCR$barcodename=paste0(name,"_",Test_TCR$barcode)
     Test_TCR$Group<-name
     AllTCR=rbind(AllTCR,Test_TCR)
  }
 
}
head(AllTCR)
```

####Load IMGT

```{r}
TRAIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/TCRgene/Gene/TRAgenelocus.csv')
TRBIMGT=read.csv('~/Main_Gao_ScanpyProject20231130/HFB_script_in_save/TCRgene/Gene/TRBlocus.csv')
TRBJIMGT=TRBIMGT[grep("TRBJ",TRBIMGT$IMGT_gene_name),]
TRAJIMGT=TRAIMGT[grep("TRAJ",TRAIMGT$IMGT_gene_name),]
```

#### Load obs

```{r}
TCRobs=read.csv("/data5/maolp/Gaofeng_All_matrix/Allcount/All_scanpyData/Data/TCRobs.csv")
TCRobs
TCRobs<-subset(TCRobs,Last_cell_type=="Naïve CD4 T")
```

```{r}
subTRA=AllTCR[AllTCR$chain=="TRA",]
subTRB=AllTCR[AllTCR$chain=="TRB",]
```


```{r}
subTRBobs<-merge(subTRB,TCRobs,by.x="barcodename",by.y="Cellname")
```
```{r}
colnames(subTRBobs)[grep("call",colnames(subTRBobs))] %>% .[grep("_1",.)]
```




```{r}
subTRBobs[,c("cdr3","IR_VDJ_1_junction_aa")]
subTRBobs2<-subTRBobs[!duplicated(subTRBobs$barcodename),]
```

### Get TCRB

```{r}
subVDJ=subTRBobs2[,c("AdjustedID","New_Body",colnames(subTRBobs2)[grep("call",colnames(subTRBobs))] %>% .[grep("_1",.)])]
# subVDJ$Week=
```
### Get TCRA
```{r}
subTRA=subTRBobs2[,c("AdjustedID","New_Body",colnames(subTRBobs2)[grep("call",colnames(subTRBobs))]
                     %>% .[grep("_1",.)])] 
# subTRA=
```




```{r}
unique(subVDJ$AdjustedID)
subVDJ$Week=substring(subVDJ$AdjustedID,2,5) %>% as.numeric()
subVDJ$Satge<-ifelse(subVDJ$Week>28,"Late","Early")
subVDJ$OrganStage<-paste0(subVDJ$Satge,"_",subVDJ$New_Body)
```

```{r}
subVDJ$TRBVJ=paste0(subVDJ$IR_VDJ_1_v_call,"_",subVDJ$IR_VDJ_1_j_call)
```

```{r}
table(subVDJ$AdjustedID,subVDJ$TRBVJ) %>% as.data.frame()
```



#----All TCR Freq----


### All percentage heatmap

```{r}
subTRBh1=subVDJ[grep("B",subVDJ$AdjustedID),]
```

```{r}
subTRBh1 = subVDJ[grep("B", subVDJ$AdjustedID), ]
TRBVJ = table(subTRBh1$TRBVJ) %>% as.data.frame() %>% arrange(desc(Freq))
TRBVJ$Per = TRBVJ$Freq / sum(TRBVJ$Freq) * 100
PBMCTRBVJ_separated <- TRBVJ %>%
  separate(Var1, into = c("Vgene", "Jgene"), sep = "_")

PBMCTRBVJ_wide <- PBMCTRBVJ_separated %>% select(-Freq) %>%
  spread(key = Vgene, value = Per)
rownames(PBMCTRBVJ_wide) <- PBMCTRBVJ_wide$Jgene
PBMCTRBVJ_wide <- PBMCTRBVJ_wide[, -1]
PBMCTRBVJ_wide[is.na(PBMCTRBVJ_wide)] <- 0
library(pheatmap)
pheatmap(
  PBMCTRBVJ_wide,
  border_color = "white",
  color = rev(colorRampPalette(
    c('#C6403D', "#f6c619", 'white', '#246BAE')
  )(100))
)
```


### All TCR VDJheatmap

#### Main Function

##### TRBIMGTsub
```{r}
TRBIMGTsub=TRBIMGT[TRBIMGT$IMGT_gene_name %in% subVDJ$IR_VDJ_1_v_call,] %>% arrange(IMGT_gene_order)
```


```{r}
library(tidyverse)
library(pheatmap)

VDJheatmap <- function(data,pattern, TRBinIMGT=TRBIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  TRBVJ = table(sub_data$TRBVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  TRBVJ$Per = TRBVJ$Freq / sum(TRBVJ$Freq)*100
  
  separated_data <- TRBVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
missing_cols <- setdiff(TRBinIMGT$IMGT_gene_name, colnames(wide_data))


for(col in missing_cols) {
  wide_data[[col]] <- 0
}
  wide_data<-wide_data[,TRBinIMGT$IMGT_gene_name]
  pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c("#C71000FF", "#edae11","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  return(wide_data)
}
```


```{r}
library(tidyverse)
library(pheatmap)

VDJheatmap_data <- function(data,pattern, TRBinIMGT=TRBIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  TRBVJ = table(sub_data$TRBVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  TRBVJ$Per = TRBVJ$Freq / sum(TRBVJ$Freq)
  
  separated_data <- TRBVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  return( wide_data)
  # wide_data<-wide_data[,TRBinIMGT$IMGT_gene_name]
  # pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c("#C71000FF", "#edae11","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  # return(wide_data)
}
```

#### Early_PBMC
```{r}
VDJheatmap(subVDJ,pattern =  "Early_PBMC",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_PBMC")
```
#### Late_PBMC no cluster
```{r}
VDJheatmap(subVDJ, pattern = "Late_PBMC",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Late_PBMC")
```

#### Late_PBMC cluster
```{r}
VDJheatmap(subVDJ, pattern = "Late_PBMC",id_col="OrganStage",cluster_col=T,cluster_row=T,title = "Late_PBMC")
```
####  Early_PBMC cluster
```{r}
VDJheatmap(subVDJ,pattern =  "Early_PBMC",id_col="OrganStage",cluster_col=T,cluster_row=,title = "Early_PBMC")
```

####  All_PBMC cluster
```{r}
VDJheatmap(subVDJ, "B",cluster_col=F,cluster_row=F,title = "PBMC")
```

#### Early_Liver
```{r}
VDJheatmap(subVDJ,pattern = "Early_Liver",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Liver")
```

#### Early_Spleen
```{r}
VDJheatmap(subVDJ,pattern = "Early_Spleen",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Spleen")
```
#### Early_Thymus
```{r}
VDJheatmap(subVDJ,pattern = "Early_Thymus",id_col="OrganStage",cluster_col=F,cluster_row=F,title = "Early_Thymus")
```


#----Per Sample TRB----

### Get Sample percentage
```{r}
subTRBh1=subVDJ[grep("B",subVDJ$AdjustedID),]
Sample_TRBVJ=table(subTRBh1$AdjustedID,subTRBh1$TRBVJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_TRBVJ)<-c("AdjustedID","VDJ","Freq")

total_freq_by_id <- Sample_TRBVJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_TRBVJ <- left_join(Sample_TRBVJ, total_freq_by_id, by = "AdjustedID")


Sample_TRBVJ <- Sample_TRBVJ %>% 
  mutate(Per = Freq / total_freq*100)

Sample_TRBVJ <- Sample_TRBVJ %>% arrange(AdjustedID)
Sample_TRBVJ$Week=substring(Sample_TRBVJ$AdjustedID,2,5) %>% as.numeric()
```
```{r}
subTRA
```




```{r}
unique(Sample_TRBVJ$AdjustedID)
```
### Sample top TRB

```{r}

library(dplyr)
library(ggplot2)

Sample_TRBVJ$Week <- as.numeric(Sample_TRBVJ$Week)
Sample_TRBVJ$Per <- as.numeric(Sample_TRBVJ$Per)


top_VDJ <- Sample_TRBVJ %>% 
  group_by(VDJ) %>%
  summarise(Total_Per = sum(Per)) %>%
  top_n(10, Total_Per) %>%
  pull(VDJ)

Sample_TRBVJ <- Sample_TRBVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% top_VDJ, VDJ, "Other"))
Sample_TRBVJ$VDJ_Top10<-factor(Sample_TRBVJ$VDJ_Top10,levels = c(as.character(top_VDJ),"Other"))
```




```{r}
table(Sample_TRBVJ$AdjustedID)
```

#### Plot: top TRB line 

```{r}

ggplot(Sample_TRBVJ, aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  geom_line() +
  scale_color_manual(values = c(rainbow(10), "gray")) +
  labs(x = "Week", y = "Per", color = "VDJ") +
  theme_classic()
```
### PerSample Cor TRBV
```{r}

PBMCTRB_cor_results <- Sample_TRBVJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
PBMCTRB_cor_results<-PBMCTRB_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(PBMCTRB_cor_results)
subset(PBMCTRB_cor_results,p.adjusted<0.1) %>% arrange(cor)
PBMCTRB_cor_results$VDJcor<-ifelse(PBMCTRB_cor_results$p.value>0.05,"NA",ifelse(PBMCTRB_cor_results$cor>0,"+","-"))
```
```{r}
subset(PBMCTRB_cor_results,p.adjusted<0.2)
```
#### Plot: top sig TRB line 



```{r}
top_VDJ=subset(PBMCTRB_cor_results,p.adjusted<0.1) %>% arrange(cor) %>% .$VDJ %>% .[-1]
Sample_TRBVJ <- Sample_TRBVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% top_VDJ, VDJ, "Other")) %>% 
mutate(VDJ_Top10=factor(VDJ_Top10,levels = c(as.character(top_VDJ),"Other")))
subset(Sample_TRBVJ,VDJ_Top10!= "Other") %>% 
ggplot() +
  geom_line( aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  scale_color_manual(values = c(rainbow(length(top_VDJ)), "gray")) +
  labs(x = "Week", y = "Percentage", color = "VDJ") +
  theme_classic()
```

```{r}
Sample_TRBVJ %>% arrange(desc(Per))
```

```{r}
# 
# PBMCTRBVJ_separated <- TRBVJ%>%
#   separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
# 
# PBMCTRBVJ_wide <- PBMCTRBVJ_separated %>%select(-Freq) %>% 
#   spread(key =Vgene, value = Per)
# rownames(PBMCTRBVJ_wide)<-PBMCTRBVJ_wide$Jgene
# PBMCTRBVJ_wide<-PBMCTRBVJ_wide[,-1]
# PBMCTRBVJ_wide[is.na(PBMCTRBVJ_wide)]<-0
```
```{r}
# library(pheatmap)
# pheatmap(PBMCTRBVJ_wide,border_color = "white",color =rev( colorRampPalette(c('#C6403D', 'white','#246BAE'))(100)))
```



```{r}
intersect(TRBIMGT$IMGT_gene_name,subVDJ$IR_VDJ_1_v_call)
```

```{r}
setdiff(subVDJ$IR_VDJ_1_v_call,TRBIMGT$IMGT_gene_name)
```





### TRBfunction Data

```{r}
library(ggplot2)
library(reshape2) # for melt function

VDJfun <- function(data,pattern, TRBinIMGT=TRBIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  TRBVJ = table(sub_data$TRBVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  TRBVJ$Per = TRBVJ$Freq / sum(TRBVJ$Freq)*100
  
  separated_data <- TRBVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(TRBinIMGT$IMGT_gene_name, colnames(wide_data))

  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  
  wide_data <- wide_data[,TRBinIMGT$IMGT_gene_name]
  
  return(wide_data)
}
```


####  TRBlongexp
```{r}
TRBFreq_exp=VDJfun(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)


TRBFreq_exp=rownames_to_column(TRBFreq_exp)
TRBlongexp <- melt(TRBFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")
colnames(TRBlongexp)<-c("Jgene","Vgene","Per")
TRBlongexp$Sample=paste0(TRBlongexp$Vgene,"_",TRBlongexp$Jgene)
TRBlongexp$VDJsig<-TRBlongexp$Sample %in% as.character(top_VDJ)


TRBlongexp$VDJsig <- ifelse(TRBlongexp$VDJsig, TRBlongexp$Sample, "Others")
```

#### PBMCTRB_cor_results
```{r}
dim(PBMCTRB_cor_results)
PBMCTRB_cor_results=PBMCTRB_cor_results %>% separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")
PBMCTRB_cor_results2=subset(PBMCTRB_cor_results,p.value<0.05&abs(cor)>0.6)
dim(TRBlongexp)
# setdiff(TRBlongexp$Sample,PBMCTRB_cor_results$VDJ)
```


#### Plot:cor percentage heatmmap 
```{r}
TRBlongexp$Jgene<-factor(TRBlongexp$Jgene,levels = rev(TRBJIMGT$IMGT_gene_name))
```
```{r}
PBMCTRB_cor_results2
```


```{r}
library(ggplot2)


PBMCTRB_cor_results2$size_factor <- cut(abs(PBMCTRB_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


TRBp1 <- ggplot() +
  geom_tile(data = TRBlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC CD4 TRB", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCTRB_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 4, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(TRBp1)
ggsave(plot = TRBp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Sup_Figure4/S25/S25A_TRB_CD4_VJcombined.pdf",width = 7,height = 7)
```
```{r}

```


#### Plot:test ggplot cluster 
```{r}
# library(reshape2)
# mat_data <- dcast(long_data, Vgene ~ Jgene, value.var = "Per")
# rownames(mat_data) <- mat_data$Vgene
# mat_data <- mat_data[, -1]
# mat_data <- as.matrix(mat_data)
# clusters_row <- hclust(dist(mat_data), method = "ave")
# clusters_col <- hclust(dist(t(mat_data)), method = "ave")
# long_data_clustered <- melt(mat_data)
# colnames(long_data_clustered) <- c("Vgene", "Jgene", "Per")
```



```{r}
# long_data_clustered$Vgene<-as.character(long_data_clustered$Vgene)
# long_data_clustered$Jgene<-as.character(long_data_clustered$Jgene)
```



```{r}
# P1 <- ggplot(long_data_clustered, aes(x =Jgene, y =  Vgene, fill = Per)) +
#   geom_tile(color = "white") +
#   scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11","white", '#246BAE')), 
#                        na.value = "white") +
#   theme_minimal() +
#   labs(title = "TRB", x = "V gene", y = "J gene") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1),
#         plot.title = element_text(hjust = 0.5)) +
#   scale_x_dendrogram(labels = T, hclust = clusters_col) +
#   scale_y_dendrogram(labels = T, hclust = clusters_row) 
# 
# print(P1)
```


```{r}
# ggplot()+geom_tile(data=PBMCTRBVJ_separated,aes(Vgene,Jgene,fill=0),color="gray52",fill="white",size=1)+geom_tile(data=PBMCTRBVJ_separated ,aes(Vgene,Jgene,value=Per,fill=Per))+ theme(axis.text.x=element_text(angle=90, hjust=1)) + scale_fill_gradientn(colours =rev(colorRampPalette(c('#C6403D', 'white','#246BAE'))(111)))+theme(axis.text.y=element_text(size = 12,color=c(rep("#246BAE",4),rep("#D0AB39",4),rep("#3FBDA7",4),rep("#C6403D",4))))+
#   guides(fill=guide_colorbar(barheight = 28))+theme_bw()+
#   theme(panel.grid = element_blank(),
#         panel.border = element_blank(),
#         axis.text.x = element_text(angle = 60,
#                                    hjust=1,vjust = 1))
```

#----Per Sample TRA----

### Get Sample percentage

```{r}
subVDJ$TRAVJ<-paste0(subVDJ$IR_VJ_1_v_call,"_",subVDJ$IR_VJ_1_j_call)
```


```{r}
subTRAh1=subVDJ[grep("B",subVDJ$AdjustedID),]
Sample_TRAVJ=table(subTRAh1$AdjustedID,subTRAh1$TRAVJ) %>% as.data.frame() %>% arrange(desc(Freq))
library(dplyr)
colnames(Sample_TRAVJ)<-c("AdjustedID","VDJ","Freq")

TRAVJ_total_freq_by_id <- Sample_TRAVJ %>% 
  group_by(AdjustedID) %>% 
  summarise(total_freq = sum(Freq))

Sample_TRAVJ <- left_join(Sample_TRAVJ, TRAVJ_total_freq_by_id, by = "AdjustedID")


Sample_TRAVJ <- Sample_TRAVJ %>% 
  mutate(Per = Freq / total_freq*100)

Sample_TRAVJ <- Sample_TRAVJ %>% arrange(AdjustedID)
Sample_TRAVJ$Week=substring(Sample_TRAVJ$AdjustedID,2,5) %>% as.numeric()
```
```{r}
library(dplyr)
library(ggplot2)

Sample_TRAVJ$Week <- as.numeric(Sample_TRAVJ$Week)
Sample_TRAVJ$Per <- as.numeric(Sample_TRAVJ$Per)


topA_VDJ <- Sample_TRAVJ %>% 
  group_by(VDJ) %>%
  summarise(Total_Per = sum(Per)) %>%
  top_n(10, Total_Per) %>%
  pull(VDJ)

Sample_TRAVJ <- Sample_TRAVJ %>% 
  mutate(VDJ_Top10 = if_else(VDJ %in% topA_VDJ, VDJ, "Other"))
Sample_TRAVJ$VDJ_Top10<-factor(Sample_TRAVJ$VDJ_Top10,levels = c(as.character(topA_VDJ),"Other"))
```

```{r}
ggplot(Sample_TRAVJ, aes(x = Week, y = Per, color = VDJ_Top10, group = VDJ)) +
  geom_line() +
  scale_color_manual(values = c(rainbow(10), "gray")) +
  labs(x = "Week", y = "Per", color = "VDJ") +
  theme_classic()
```
#### corTRA
```{r}
PBMCTRA_cor_results <- Sample_TRAVJ %>%
  group_by(VDJ) %>%
  summarise(cor = cor.test(Week, Per, method = "spearman")$estimate,
            p.value = cor.test(Week, Per, method = "spearman")$p.value)
PBMCTRA_cor_results<-PBMCTRA_cor_results %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH"))%>% arrange(cor)

print(PBMCTRA_cor_results)
subset(PBMCTRA_cor_results,p.value<0.05) %>% arrange(cor)
PBMCTRA_cor_results$VDJcor<-ifelse(PBMCTRA_cor_results$p.value>0.05,"NA",ifelse(PBMCTRA_cor_results$cor>0,"+","-"))
```

### TRAfunction Data

```{r}
library(ggplot2)
library(reshape2) # for melt function

TRAVDJfun <- function(data,pattern, TRAinIMGT=TRAIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  TRAVJ = table(sub_data$TRAVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  TRAVJ$Per = TRAVJ$Freq / sum(TRAVJ$Freq)*100
  
  separated_data <- TRAVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(TRAinIMGT$IMGT_gene_name, colnames(wide_data))
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  
  wide_data <- wide_data[,TRAinIMGT$IMGT_gene_name]
  
  return(wide_data)
}
```

```{r}
TRAIMGTsub=TRAIMGT[TRAIMGT$IMGT_gene_name %in% subVDJ$IR_VJ_1_v_call,] %>% arrange(IMGT_locus_gene_order)
```
####  TRAlongexp
```{r}
TRAFreq_exp=TRAVDJfun(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)


TRAFreq_exp=rownames_to_column(TRAFreq_exp)
TRAlongexp <- melt(TRAFreq_exp, varnames = c("rowname", "Vgene"), value.name = "Per")
colnames(TRAlongexp)<-c("Jgene","Vgene","Per")
TRAlongexp$Sample=paste0(TRAlongexp$Vgene,"_",TRAlongexp$Jgene)
TRAlongexp$VDJsig<-TRAlongexp$Sample %in% as.character(top_VDJ)


TRAlongexp$VDJsig <- ifelse(TRAlongexp$VDJsig, TRAlongexp$Sample, "Others")
```

#### PBMCTRA_cor_results
```{r}
dim(PBMCTRA_cor_results)
PBMCTRA_cor_results=PBMCTRA_cor_results %>% separate(VDJ, into = c("Vgene", "Jgene"), sep = "_")
PBMCTRA_cor_results2=subset(PBMCTRA_cor_results,p.value<0.05&abs(cor)>0.6)
dim(TRAlongexp)
# setdiff(TRAlongexp$Sample,PBMCTRA_cor_results$VDJ)
```


#### Plot:cor percentage heatmmap 
```{r}
TRAlongexp$Jgene<-factor(TRAlongexp$Jgene,levels = rev(TRAJIMGT$IMGT_gene_name))
```
```{r}
PBMCTRA_cor_results2
```
```{r}
system("mkdir -p /data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Sup_Figure4/S25")
```

```{r}
library(ggplot2)


PBMCTRA_cor_results2$size_factor <- cut(abs(PBMCTRA_cor_results2$cor),
                                        breaks = c(-Inf, 0.5, 0.6,0.7,0.8, Inf),
                                        labels = c("0-0.5","0.5-0.6", "0.6-0.7","0.7-0.8", "0.8-1.0"))


TRAp1 <- ggplot() +
  geom_tile(data = TRAlongexp, aes(x = Vgene, y = Jgene, fill = Per), color = "white") +
  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11", "white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC CD4 TRA", x = "V gene", y = "J gene", fill = "Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(data = PBMCTRA_cor_results2, 
            aes(x = Vgene, y = Jgene, label = VDJcor, color = VDJcor, size = size_factor)) +
  scale_color_manual(values = c("#3D3B25FF", "#8F1336")) +
  scale_size_manual(values = c("0-0.5" = 5, "0.6-0.7" = 5, "0.7-0.8"= 6,"0.8-0.9"= 7,"0.8-1.0"=8))


print(TRAp1)
  ggsave(plot = TRAp1,"/data5/maolp/Main_Gao_ScanpyProject20231130/HFB_Figure_Plot/Sup_Figure4/S25/S25A_TRA_CD4_VJcombined.pdf",width = 7,height = 7)
```

```{r}
TRAVDJheatmap <- function(data,pattern, TRAinIMGT=TRAIMGTsub ,id_col="AdjustedID", cluster_col = TRUE, cluster_row = TRUE, title = "Organ") {
  sub_data = data[grep(pattern, data[[id_col]]),]
  TRAVJ = table(sub_data$TRAVJ) %>% as.data.frame() %>% arrange(desc(Freq))
  TRAVJ$Per = TRAVJ$Freq / sum(TRAVJ$Freq)*100
  
  separated_data <- TRAVJ %>%
    separate(Var1, into = c("Vgene", "Jgene"), sep = "_")
  
  wide_data <- separated_data %>% select(-Freq) %>%
    spread(key = Vgene, value = Per)
  
  rownames(wide_data) <- wide_data$Jgene
  wide_data <- wide_data[, -1]
  wide_data[is.na(wide_data)] <- 0
  missing_cols <- setdiff(TRAinIMGT$IMGT_gene_name, colnames(wide_data))
  
  
  for(col in missing_cols) {
    wide_data[[col]] <- 0
  }
  wide_data<-wide_data[,TRAinIMGT$IMGT_gene_name]
  pheatmap(wide_data, border_color = "white", color = rev(colorRampPalette(c("#C71000FF", "#edae11","white", '#246BAE'))(100)), cluster_cols = cluster_col, cluster_rows = cluster_row, main = title)
  return(wide_data)
}
```

```{r}
TRAVDJheatmap(data = subVDJ, pattern = "B", id_col = "AdjustedID",title = "PBMC",cluster_col = F)
```

# ----TRB----

```{r}

subTRBvdj=subTRBobs2[, c("AdjustedID", "New_Body", colnames(subTRBobs2)[grep("VDJ", colnames(subTRBobs2))])]
subTRBvdj$TRBvj=paste0(subTRBvdj$IR_VDJ_1_v_call,"_",subTRBvdj$IR_VDJ_1_j_call)
subTRBvdj$aalength=nchar(subTRBvdj$IR_VDJ_1_junction_aa)
subTRBvdjlen=table(subTRBvdj$TRBvj,subTRBvdj$aalength) %>% as.data.frame()
```

```{r}
# subset(subTRBvdjlen,Len==17&Freq>median((subset(subTRBvdjlen,Len==17&Freq>0)$Freq)))
# max(subset(subTRBvdjlen,Len==17&Freq>0)$Freq)
```
```{r}
PBMCTRB_cor_results2$VJgene=paste0(PBMCTRB_cor_results2$Vgene,"_",PBMCTRB_cor_results2$Jgene)
# intersect(PBMCTRB_cor_results2$VJgene,subset(subTRBvdjlen,Len==17&Freq>median((subset(subTRBvdjlen,Len==17&Freq>0)$Freq)))$VDJ)
```


```{r}
colnames(subTRBvdjlen)<-c("VDJ","Len","Freq")
subTRBvdjlen$Percentage=subTRBvdjlen$Freq/sum(subTRBvdjlen$Freq)*100

P2=ggplot(subTRBvdjlen)+geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11","white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC TRA", x = "V gene", y = "J gene",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
P2
```

```{r}

subTRBvdjlen[subTRBvdjlen$VDJ %in% intersect(PBMCTRB_cor_results2$VJgene,subset(subTRBvdjlen,Len==17&Freq>median((subset(subTRBvdjlen,Len==17&Freq>0)$Freq)))$VDJ),] %>%
  ggplot() +geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11","white", '#246BAE')), 
                       na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC TRB", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}
# subTRBvdjlen=table(subTRBvdj$TRBvj,subTRBvdj$aalength) %>% as.data.frame()
```


```{r}
PBMCTRB_cor_results2$VDJgene=paste0(PBMCTRB_cor_results2$Vgene,"_",PBMCTRB_cor_results2$Jgene)
TRBlongexpsig=merge(TRBlongexp,PBMCTRB_cor_results2,by.x="Sample",by.y="VDJgene")
subset(TRBlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10)
# max_five <- sort(TRBlongexp$Per, decreasing = TRUE) %>% head( 10)

subset(TRBlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subset(TRBlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subTRBsortgene=subset(TRBlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor) %>% .$Sample
subTRBvdjlen[subTRBvdjlen$VDJ %in%subTRBsortgene ,] %>% mutate(VDJ=factor(VDJ,levels=rev(subTRBsortgene)))%>%
  ggplot() +geom_tile(aes(Len,VDJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11","white", '#246BAE')), 
                                                                                        na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC TRB", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```
```{r}
# subTRBvdjlenmax
```

### Plot:TRB All len
```{r}
subTRBvdjlenmax <- subTRBvdjlen %>%
  group_by(VDJ) %>%
  mutate(Percentage_max = max(Percentage)) %>%
  filter(Percentage == Percentage_max)
m1=merge(subTRBvdjlenmax,TRBlongexpsig,by.x="VDJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VDJ %in% subTRBsortgene,subTRBsortgene,"Others")) 
m1<-m1[!duplicated(m1$VDJ),]
length(unique(m1$VDJ))
m1$VDJcol <- subTRBsortgene[match(m1$VDJ, subTRBsortgene)]
m1$VDJcol[is.na(m1$VDJcol)] <- "Others TRB V-J"
table(m1$VDJcol)
setdiff( subTRBsortgene,m1$VDJ )

library(ggsci)

P3TRBplot=m1%>% mutate(Rcor=ifelse(cor<0,"R<0","R>0")) %>% mutate(VDJcol=factor(VDJcol,levels=c(subTRBsortgene,"Others TRB V-J"))) %>% 
  ggplot() +
  geom_point(aes(Len,cor,size=Per,color=VDJcol,shape=Rcor)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC CD4 TRB", x = "TRB V-J CDR3 max len", y = " Rcor",size="Percentage in TRB V-J(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),aspect.ratio = 0.6)+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))+scale_color_futurama()
P3TRBplot
```
```{r}
# ggsave(plot=P3TRBplot,"Figure/P3TRBplot.pdf",width = 8,height = 6)
```
### Plot:TRB All len
```{r}
merge(subTRBvdjlen,TRBlongexpsig,by.x="VDJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VDJ %in% subTRBsortgene,subTRBsortgene,"Others")) %>% mutate(Percentage2=ifelse(cor <0,-Percentage,Percentage)) %>% 
  ggplot() +
  geom_point(aes(Len,Percentage2,color=VDJcol)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC TRB", x = "VDJ CDR3 max len", y = "Percentage(%)",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))
```


#----TRA----
```{r}
subTRAvdj=subTRBobs2[, c("AdjustedID", "New_Body", colnames(subTRBobs2)[grep("VJ", colnames(subTRBobs2))])]
subTRAvdj$TRAvj=paste0(subTRAvdj$IR_VJ_1_v_call,"_",subTRAvdj$IR_VJ_1_j_call)
subTRAvdj$aalength=nchar(subTRAvdj$IR_VJ_1_junction_aa)
subTRAvdjlen=table(subTRAvdj$TRAvj,subTRAvdj$aalength) %>% as.data.frame()
max(subTRAvdjlen$Percentage)
colnames(subTRAvdjlen)<-c("VJ","Len","Freq")
subTRAvdjlen$Percentage=subTRAvdjlen$Freq/sum(subTRAvdjlen$Freq)*100
```
### Plot:TRA All len
```{r}
PlenTRA=ggplot(subTRAvdjlen)+geom_tile(aes(Len,VJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c('#C6403D', "#f6c619", '#246BAE')), 
                                                                                                     na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC TRA", x = "CDR3 Length", y = "J gene",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank())
PlenTRA
```

```{r}
# ggsave(plot=PlenTRA,"Figure/PlenTRA.pdf")
```

```{r}
PBMCTRA_cor_results2$VJgene[1:10]
```
```{r}
PBMCTRA_cor_results2$VJgene=paste0(PBMCTRA_cor_results2$Vgene,"_",PBMCTRA_cor_results2$Jgene)
TRAlongexpsig=merge(TRAlongexp,PBMCTRA_cor_results2,by.x="Sample",by.y="VJgene")
subset(TRAlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10)
max_five <- sort(TRAlongexp$Per, decreasing = TRUE) %>% head( 10)

subset(TRAlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```

```{r}
subset(TRAlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor)
```
### Plot:TRA ten len
```{r}
subTRAsortgene=subset(TRAlongexpsig,p.value<0.05) %>% arrange(desc(Per)) %>% head( 10) %>%  arrange(cor) %>% .$Sample
subTRAvdjlen[subTRAvdjlen$VJ %in%subTRAsortgene ,] %>% mutate(VJ=factor(VJ,levels=rev(subTRAsortgene)))%>%
  ggplot() +geom_tile(aes(Len,VJ,fill=Percentage),color="white")+  scale_fill_gradientn(colors = rev(c("#C71000FF", "#edae11","white", '#246BAE')), 
                                                                                         na.value = "white") +
  theme_minimal() +
  labs(title = "PBMC TRA", x = "", y = " ",fill="Percentage(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```


```{r}
subTRAvdjlenmax <- subTRAvdjlen %>%
  group_by(VJ) %>%
  mutate(Percentage_max = max(Percentage)) %>%
  filter(Percentage == Percentage_max)
m1=merge(subTRAvdjlenmax,TRAlongexpsig,by.x="VJ",by.y="Sample") %>% mutate(VDJcol=ifelse(VJ %in% subTRAsortgene,subTRAsortgene,"Others")) 
m1<-m1[!duplicated(m1$VJ),]
length(unique(m1$VJ))
m1$VDJcol <- subTRAsortgene[match(m1$VJ, subTRAsortgene)]
m1$VDJcol[is.na(m1$VDJcol)] <- "Others TRA V-J"
table(m1$VDJcol)
setdiff( subTRAsortgene,m1$VJ )

library(ggsci)

P3TRAplot=m1%>% mutate(Rcor=ifelse(cor<0,"R<0","R>0")) %>% mutate(VDJcol=factor(VDJcol,levels=c(subTRAsortgene,"Others TRA V-J"))) %>% 
  ggplot() +
  geom_point(aes(Len,cor,size=Per,color=VDJcol,shape=Rcor)) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "black") +
  theme_classic() +
  labs(title = "PBMC CD4 TRA", x = "TRA V-J CDR3 max len", y = " Rcor",size="Percentage in TRA V-J(%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),aspect.ratio = 0.6)+scale_y_continuous(limits = c(-1, 1),breaks = seq(-1,1,0.2))+scale_color_futurama()
P3TRAplot
```

```{r}
# ggsave(plot=P3TRAplot,"Figure/P3TRAplot.pdf")
```

### save and load

```{r}
### 2023-10-28 23:30:58
save.image("DATA/CD4_VDJcombined.RData")
```

